---
title: "Hands-on_Ex4"
author: "Chia Yong Soon"
date: "09 Dec 2023"
date-modified: "last-modified"
---

## 13  Calibrating Hedonic Pricing Model for Private Highrise Property with GWR Method

### 13.1 Overview
Geographically weighted regression (GWR) is a spatial statistical technique that takes non-stationary variables into consideration (e.g., climate; demographic factors; physical environment characteristics) and models the local relationships between these independent variables and an outcome of interest (also known as dependent variable). In this hands-on exercise, you will learn how to build hedonic pricing models by using GWR methods. The dependent variable is the resale prices of condominium in 2015. The independent variables are divided into either structural and locational.

### 13.2 The Data
Two data sets will be used in this model building exercise, they are:

URA Master Plan subzone boundary in shapefile format (i.e. MP14_SUBZONE_WEB_PL)
condo_resale_2015 in csv format (i.e. condo_resale_2015.csv)
13.3 Getting Started
Before we get started, it is important for us to install the necessary R packages into R and launch these R packages into R environment.

The R packages needed for this exercise are as follows:

R package for building OLS and performing diagnostics tests
olsrr
R package for calibrating geographical weighted family of models
GWmodel
R package for multivariate data visualisation and analysis
corrplot
Spatial data handling
sf
Attribute data handling
tidyverse, especially readr, ggplot2 and dplyr
Choropleth mapping
tmap
The code chunks below installs and launches these R packages into R environment.

```{r}
pacman::p_load(olsrr, corrplot, ggpubr, sf, spdep, GWmodel, tmap, tidyverse, gtsummary)
```

### 13.4 A short note about GWmodel
GWmodel package provides a collection of localised spatial statistical methods, namely: GW summary statistics, GW principal components analysis, GW discriminant analysis and various forms of GW regression; some of which are provided in basic and robust (outlier resistant) forms. Commonly, outputs or parameters of the GWmodel are mapped to provide a useful exploratory tool, which can often precede (and direct) a more traditional or sophisticated statistical analysis.

### 13.5 Geospatial Data Wrangling
#### 13.5.1 Importing geospatial data
The geospatial data used in this hands-on exercise is called MP14_SUBZONE_WEB_PL. It is in ESRI shapefile format. The shapefile consists of URA Master Plan 2014’s planning subzone boundaries. Polygon features are used to represent these geographic boundaries. The GIS data is in svy21 projected coordinates systems.

The code chunk below is used to import MP_SUBZONE_WEB_PL shapefile by using st_read() of sf packages.

```{r}
mpsz = st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

#### 13.5.2 Updating CRS information
The code chunk below updates the newly imported mpsz with the correct ESPG code (i.e. 3414)

```{r}
mpsz_svy21 <- st_transform(mpsz, 3414)
```

After transforming the projection metadata, you can varify the projection of the newly transformed mpsz_svy21 by using st_crs() of sf package.

The code chunk below will be used to varify the newly transformed mpsz_svy21.

```{r}
st_crs(mpsz_svy21)
```


Notice that the EPSG: is indicated as 3414 now.

Next, you will reveal the extent of mpsz_svy21 by using st_bbox() of sf package.
```{r}
st_bbox(mpsz_svy21) #view extent
```

### 13.6 Aspatial Data Wrangling
#### 13.6.1 Importing the aspatial data
The condo_resale_2015 is in csv file format. The codes chunk below uses read_csv() function of readr package to import condo_resale_2015 into R as a tibble data frame called condo_resale.

```{r}
condo_resale = read_csv("data/aspatial/Condo_resale_2015.csv")
```

After importing the data file into R, it is important for us to examine if the data file has been imported correctly.

The codes chunks below uses glimpse() to display the data structure of will do the job.

```{r}
glimpse(condo_resale)
```

```{r}
head(condo_resale$LONGITUDE) #see the data in XCOORD column
```

```{r}
head(condo_resale$LATITUDE) #see the data in YCOORD column
```

Next, summary() of base R is used to display the summary statistics of cond_resale tibble data frame.

```{r}
summary(condo_resale)
```

#### 13.6.2 Converting aspatial data frame into a sf object
Currently, the condo_resale tibble data frame is aspatial. We will convert it to a sf object. The code chunk below converts condo_resale data frame into a simple feature data frame by using st_as_sf() of sf packages.

```{r}
condo_resale.sf <- st_as_sf(condo_resale,
                            coords = c("LONGITUDE", "LATITUDE"),
                            crs=4326) %>%
  st_transform(crs=3414)
```


Notice that st_transform() of sf package is used to convert the coordinates from wgs84 (i.e. crs:4326) to svy21 (i.e. crs=3414).

Next, head() is used to list the content of condo_resale.sf object.

```{r}
head(condo_resale.sf)
```