---
title: "In-class_Ex2"
author: "Yong Soon"
date: "25 Nov 2023"
date-modified: "date-modified"
---

## Getting started

### Installing and Loading the R Packages

Four R packages will be used for this in-class exercise, theu are sf, sfdep, tmap, tidyverse.

::: {style="font-size: 1.5em"}
```{r}
pacman::p_load(sf, sfdep, tmap, tidyverse, knitr, plotly)
# knitr is to generate HTML table
# tidyverse consists of many different packages
```
:::

### The Data

For the purpose of this in-class exercise, the Hunan data sets will be used. Theree are two data sets in this use case, they are:

- Hunan, a geospatial data set in ESRI sharefile format, and
- Hunan_2012, an attribute data set in csv format

#### Importing geospatial data

```{r}
Hunan = st_read(dsn = "data/geospatial", 
                         layer = "Hunan")
```


#### Importing attribute table

```{r}
Hunan_2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```


#### Combining both dataframe by using left join


#### Plotting a choropleth map



### Deriving contiguity Spatial Weights



### Deriving contiguity weights Queen's method



### Distance-based Weights

