---
title: "In-class Exercise 1"
date: "18 November 2023"
format: html
editor: visual
---
## 1 Getting Started

In this hands-on exercise, two R packages will be used. They are:

sf for importing, managing, and processing geospatial data, and tidyverse for performing data science tasks such as importing, wrangling and visualising data. Tidyverse consists of a family of R packages. In this hands-on exercise, the following packages will be used:

readr for importing csv data, readxl for importing Excel worksheet, tidyr for manipulating data, dplyr for transforming data, and ggplot2 for visualising data

Type the following code chunk.

```{r}
pacman::p_load(sf, tidyverse)
```

What we can learn from the code chunk above:

p_load function pf pacman package is used to install and load sf and tidyverse pacages into R environment.

## 1.1 Importing Geospatial Data

In this section, you will learn how to import the following geospatial data into R by using st_read() of sf package:

MPSZ-2019, a polygon feature layer in ESRI shapefile format, BusStop, a line feature layer in ESRI shapefile format.

#### 1.1.1 Importing polygon feature data in shapefile format

The code chunk below uses st_read() function of sf package to import MPSZ-2019 shapefile into R as a polygon feature data frame. Note that when the input geospatial data is in shapefile format, two arguments will be used, namely: dsn to define the data path and layer to provide the shapefile name. Also note that no extension such as .shp, .dbf, .prj and .shx are needed.

```{r}
mpsz = st_read(dsn = "data/geospatial", 
                  layer = "MPSZ-2019")
```


#### 1.1.2 Importing polyline feature data in shapefile form

The code chunk below uses st_read() function of sf package to import BusStop shapefile into R as line feature data frame.

```{r}
cyclingpath = st_read(dsn = "data/geospatial", 
                         layer = "BusStop")
```

The message above reveals that there are a total of 5161 features and 3 fields in BusStop linestring feature data frame and it is in svy21 projected coordinates system too.

